Blaming the Typeless: \\ Scalable, Human-Centric Python Localization

### 2018-08-18 paper/main.tex
paper updates

@@

\bibliographystyle{abbrv}
\bibliography{nanomaly,sw}

>>>>

\section{Approach, in more detail}

We follow the overall strategy of \cite{learning-to-blame} and represent each
buggy program by a Bag of Abstracted Terms. In our setting, a `term' means either
a statment or an expression. For each term in the program we compute the following features:

\subsection{Dynamic features}

We first convert each buggy program to an equivalent program that is closer
to ANF, then run it through the PythonTutor backend. This backend is based on the
Python debigging library BDB; it returns to us a trace of the program along with
the state of the heap at each execution step. We use this information to compute
the following features for each term:

\emph{Type} The possible values for the Type feature include all the basic Python
types (such as int and tuple) as well as a few special values:
\begin{itemize}
    \item IsStatement - Given to all statements, since our terms include both
    statments and expressions but only expressions actually have types.
    \item Unknown - Given to expressions that are never evaluated in the dynamic
    trace.
    \item Multiple - Given to expressions which are evaluated multiple times in
    a trace and do not always have the same type.
\end{itemize}

\emph{Slice} (TODO - Leon can you expand this? What did you base our slicing on?)
The goal of this feature is to help eliminate terms that cannot be the source of
the crash. These include lines that are never run, lines that run successfully
and don't affect any variables that are used later, and so on. More specifically, we
build up a dependency graph of all lines in the program trace. Edges in the graph
represent either def-use dependence or control-flow dependence. Then ideally,
any node in the graph that is not a transitive dependency of the line that crashes
can't be responsible for the crash. However dynamic slicing is not perfect (TODO
include example?), so unlike in prior work, we do not treat membership in the slice
as a hard constraint.

\emph{Crash location} As discussed later in the evaluation, the precise term that
causes Python to crash is in fact the one the user decides to change about 30\% of
the time, so we use it as a feature (and expect it to be a high-signal one, too.)

\emph{Exception type} The type of error thrown by the program is useful context for
finding the error. For example, if asked how likely a random `0` is of being a bug, you'd give
a higher answer if you knew that the program eventually crashes with a division by
zero exception. Thus we attach the exception type as a feature to \emph{every}
vector derived from that program.

\subsection{Contextual features}

The approach of representing a program by a Bag of Abstracted Terms, if implemented
naively, runs into the problem that the terms lose their context in the program
as a whole. For example, the features for the term `0` include its type (int) and
syntactic description (int literal), but there is no way that's enough to decide
if it's innocuous, like the 0 in `x = 0`, or clearly the source of a bug, like the
0 in `x / 0`. To restore some of this context, we introduce contextual features.
This involves duplicating all our other features four times - once for the parent
and once each for a term's first three children in the AST. So since `x / 0` has
the feature SyntacticForm = DivisionOp, the `0` of `x / 0` has the feature
ParentSyntacticForm = DivisionOp. Similarly, since `0` has SyntacticForm = IntLiteral,
`x / 0` has Child1SyntacticForm = IntLiteral. In order to make all feature vectors
the same length, nodes that have no parent or fewer than three children have the
relevant features filled out with a special NotApplicable value.

\subsection{Static features}

\emph{Syntactic form} The first feature we consider is what kind of node the term
is in the AST. For statements this has values like Return or Import; for expressions
Variable or Application.

\emph{Size} We also count the number of descendent nodes in the subtree rooted at that node.

\subsection{Machine learning}

\emph{Pairing programs} For each PythonTutor user, we look at the sequence of
programs they attempted to run. Each program that crashes is paired with the next
program by that user which does not crash. Many programs do not make it to the
end of our data pipeline; here are the most common reasons:
\begin{itemize}
    \item The user submitted only programs that crash or only programs that do
    not crash, so we can't create any pairs for that user.
    \item The program is out of scope for PythonTutor. For
    example, the program imports forbidden libraries (like os), or the program
    runs in over 1000 execution steps.
    \item The proportion of the program that has changed between the buggy and
    fixed versions is more than a standard deviation above the mean.
\end{itemize}
We end up with \textasciitilde270,000 usable pairs of programs.

\emph{Training methodology} At this point we have feature vectors describing every term of every program in
our dataset. The vectors from a random 30\% of the program pairs are set aside for testing and
the other 70\% are used for training. Within the training data, we use 3-fold
cross-validation to select the best parameters for our models, before training
using those parameters on the full training set and then scoring the models
using the testing set. Each trained model takes in any vector representing a term
in a buggy program, and returns a confidence score representing how likely it is that
that term was one of the terms changed between the fixed and buggy versions.
Thus we can treat the model as providing a \emph{ranking} over all terms by
confidence, where the top result is the one that the model deems most likely
to be changed. For
a given k, we score the model based on Top-k accuracy: for what proportion of the
programs is a correct answer (i.e. a term that is actually changed) present in the
top k results. We retrain the models for each value of k, since a model optimized
to produce the single best result might be different from one optimized to get a
good result into the top three.

In each case the most effective model we trained was a decision tree. (The other
models we tried were Logistic Regression and MLP. Yaoyuan - anything to add to
this section?)

@@

\bibliographystyle{abbrv}
\bibliography{nanomaly,sw,temp}
